Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2024-03-05 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
cotton conference changes done

---
title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
editor: source
---

@@

# Flowchart

```{r}
grViz("digraph {

# initiate graph
graph [layout = dot]

# global node settings
node [shape = rectangle]

# label nodes
data1 [label = '@@1']
data2 [label = '@@2']
ineligible [label = '@@3']
demo  [label = '@@4']
highna  [label = '@@5']
highcl  [label = '@@6']
highnacl  [label = '@@7']
review [label = '@@8']
analysis [label = '@@9']

# edge definitions with the node IDs
data1  -> data2
data2 -> ineligible
ineligible -> {demo,highna,highcl,highnacl,review}
data2  -> analysis
highnacl -> analysis [style=invis]
}
[1]: paste0('Sweat tests performed \\n ','n = ',n1)
[2]: paste0('Unique participants \\n ','n = ',n2)
[3]: paste0('Ineligible \\n ','n = ',n2-nrow(t1_df))
[4]: paste0('No demographics \\n ','n = ',length(no_dem))
[5]: paste0('Na > 160 mmol/L \\n ','n = ',length(high_na))
[6]: paste0('Cl > 160 mmol/L \\n ','n = ',length(high_cl))
[7]: paste0('Na:Cl > 2 \\n ','n = ',length(high_ratio))
[8]: paste0('Other \\n ','n = ',1)
[9]: paste0('Analysis cohort \\n ','n = ',nrow(t1_df))
")
```

@@

@@

plot(roc_cf, print.auc = TRUE, col = "blue")
plot(roc_non_cf, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
```

The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs. CF, and the green is the ROC curve for distinguishing between non-CF vs. CFTR-RD or CRMS. The optimal threshold for distinguishing between CFTR-RD or CRMS vs. CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index. The optimal threshold for distinguishing between non-CF vs. CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index. The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.

@@

Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and green indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.

```{r warning=FALSE}
ggplot(
  t1_df[!is.na(t1_df$`Diagnosis`), ],
  aes(log(`Na/Cl Ratio`), fill = `Diagnosis`)
) +
  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
  geom_vline(xintercept = log(coords_non_cf$threshold), color = "green") +
  ylab("Count") +
  xlab("ln(Na/Cl Ratio)") +
  theme_bw()
```

@@

plot(roc_cf, print.auc = TRUE, col = "blue")
plot(roc_non_cf, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
```

The blue curve shows the ROC for distinguishing between low vs. high risk, and the green is the ROC curve for distinguishing between non-CF vs. low risk. The optimal threshold for distinguishing between low vs. high risk was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index. The optimal threshold for distinguishing between non-CF vs. low risk was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index. The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.

@@

Vertical blue line indicates the threshold for low vs. high risk and green indicates the threshold for non-CF vs. low risk. Na/Cl ratio was log-transformed for easier visualization.

```{r warning=FALSE}
ggplot(
  t1_df[!is.na(t1_df$`Genotype Risk`), ],
  aes(log(`Na/Cl Ratio`), fill = `Genotype Risk`)
) +
  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
  geom_vline(xintercept = log(coords_non_cf$threshold), color = "green") +
  ylab("Count") +
  xlab("ln(Na/Cl Ratio)") +
  theme_bw()
```

@@

plot(roc_cf, print.auc = TRUE, col = "blue")
plot(roc_non_cf, print.auc = TRUE, col = "green", print.auc.y = .4, add = TRUE)
```

The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs. CF, and the green is the ROC curve for distinguishing between non-CF vs. CFTR-RD or CRMS. The optimal threshold for distinguishing between CFTR-RD or CRMS vs. CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index. The optimal threshold for distinguishing between non-CF vs. CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index. The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.

>>>>

---
title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 10
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
editor: source
---

@@

<!-- # Flowchart -->

<!-- ```{r} -->
<!-- grViz("digraph { -->

<!-- # initiate graph -->
<!-- graph [layout = dot] -->

<!-- # global node settings -->
<!-- node [shape = rectangle] -->

<!-- # label nodes -->
<!-- data1 [label = '@@1'] -->
<!-- data2 [label = '@@2'] -->
<!-- ineligible [label = '@@3'] -->
<!-- demo  [label = '@@4'] -->
<!-- highna  [label = '@@5'] -->
<!-- highcl  [label = '@@6'] -->
<!-- highnacl  [label = '@@7'] -->
<!-- review [label = '@@8'] -->
<!-- analysis [label = '@@9'] -->

<!-- # edge definitions with the node IDs -->
<!-- data1  -> data2 -->
<!-- data2 -> ineligible -->
<!-- ineligible -> {demo,highna,highcl,highnacl,review} -->
<!-- data2  -> analysis -->
<!-- highnacl -> analysis [style=invis] -->
<!-- } -->
<!-- [1]: paste0('Sweat tests performed \\n ','n = ',n1) -->
<!-- [2]: paste0('Unique participants \\n ','n = ',n2) -->
<!-- [3]: paste0('Ineligible \\n ','n = ',n2-nrow(t1_df)) -->
<!-- [4]: paste0('No demographics \\n ','n = ',length(no_dem)) -->
<!-- [5]: paste0('Na > 160 mmol/L \\n ','n = ',length(high_na)) -->
<!-- [6]: paste0('Cl > 160 mmol/L \\n ','n = ',length(high_cl)) -->
<!-- [7]: paste0('Na:Cl > 2 \\n ','n = ',length(high_ratio)) -->
<!-- [8]: paste0('Other \\n ','n = ',1) -->
<!-- [9]: paste0('Analysis cohort \\n ','n = ',nrow(t1_df)) -->
<!-- ") -->
<!-- ``` -->

@@

##### Na/Cl Ratio

@@

plot(roc_cf, print.auc = TRUE, col = "blue")
plot(roc_non_cf, print.auc = TRUE, col = "red")
```

The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs. CF, and the red is the ROC curve for distinguishing between non-CF vs. CFTR-RD or CRMS. The optimal threshold for distinguishing between CFTR-RD or CRMS vs. CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index. The optimal threshold for distinguishing between non-CF vs. CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index. The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.

##### Cl Alone

```{r}
roc_cf_cl <- roc(
  response = t1_df$`Diagnosis`,
  predictor = t1_df$Cl,
  levels = c("CFTR-RD or CRMS", "Cystic Fibrosis"),
  quiet = T
)
coords_cf <- round(coords(roc_cf_cl, "best", best.method = "youden"), 3)

roc_non_cf_cl <- roc(
  response = t1_df$`Diagnosis`,
  predictor = t1_df$Cl,
  levels = c("Non-CF", "CFTR-RD or CRMS"),
  quiet = T
)
coords_non_cf <- round(coords(roc_non_cf_cl, "best", best.method = "youden"), 3)

plot(roc_cf_cl, print.auc = TRUE, col = "blue")
plot(roc_non_cf_cl, print.auc = TRUE, col = "red")
```

The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs. CF, and the red is the ROC curve for distinguishing between non-CF vs. CFTR-RD or CRMS. The optimal threshold for distinguishing between CFTR-RD or CRMS vs. CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index. The optimal threshold for distinguishing between non-CF vs. CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index. The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.

##### ROC Comparison

###### CFTR-RD or CRMS vs. CF

```{r}
roc.test(roc_cf,roc_cf_cl)
```

###### Non-CF vs. CFTR-RD or CRMS

```{r}
roc.test(roc_non_cf,roc_non_cf_cl)
```

@@

Vertical blue line indicates the threshold for CFTR-RD or CRMS vs. CF and red indicates the threshold for non-CF vs. CFTR-RD or CRMS. Na/Cl ratio was log-transformed for easier visualization.

```{r warning=FALSE}
ggplot(
  t1_df[!is.na(t1_df$`Diagnosis`), ],
  aes(log(`Na/Cl Ratio`), fill = `Diagnosis`)
) +
  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
  ylab("Count") +
  xlab("ln(Na/Cl Ratio)") +
  theme_bw()
```

@@

plot(roc_cf, print.auc = TRUE, col = "blue")
plot(roc_non_cf, print.auc = TRUE, col = "red", print.auc.y = .4, add = TRUE)
```

The blue curve shows the ROC for distinguishing between low vs. high risk, and the red is the ROC curve for distinguishing between non-CF vs. low risk. The optimal threshold for distinguishing between low vs. high risk was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index. The optimal threshold for distinguishing between non-CF vs. low risk was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index. The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.

@@

Vertical blue line indicates the threshold for low vs. high risk and red indicates the threshold for non-CF vs. low risk. Na/Cl ratio was log-transformed for easier visualization.

```{r warning=FALSE}
ggplot(
  t1_df[!is.na(t1_df$`Genotype Risk`), ],
  aes(log(`Na/Cl Ratio`), fill = `Genotype Risk`)
) +
  geom_histogram(alpha = 0.3, position = "identity", bins = 50) +
  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
  geom_vline(xintercept = log(coords_non_cf$threshold), color = "red") +
  ylab("Count") +
  xlab("ln(Na/Cl Ratio)") +
  theme_bw()
```

@@

plot(roc_cf, print.auc = TRUE, col = "blue")
plot(roc_non_cf, print.auc = TRUE, col = "red", print.auc.y = .4, add = TRUE)
```

The blue curve shows the ROC for distinguishing between CFTR-RD or CRMS vs. CF, and the red is the ROC curve for distinguishing between non-CF vs. CFTR-RD or CRMS. The optimal threshold for distinguishing between CFTR-RD or CRMS vs. CF was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index. The optimal threshold for distinguishing between non-CF vs. CFTR-RD or CRMS was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index. The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.
